---
title: "DS311 - R Lab Assignment"
author: "Vickie Lam"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
ncol(mtcars)
nrow(mtcars)

# Answer:
print("There are total of ___11__ variables and __32___ observations in this data set.")
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)

# Answer:
print("There are ___7__ discrete variables and ___4__ continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!

m <- mean(mtcars$mpg)
v <- var(mtcars$mpg)
s <- sd(mtcars$mpg)



print(paste("The average of Mile Per Gallon from this data set is ", m , " with variance ", v , " and standard deviation", s , "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!

avg_mpg_cyl <- aggregate(mpg ~ cyl, data = mtcars, mean)
avg_mpg_cyl

sd_mpg_gear <- aggregate(mpg ~ gear, data = mtcars, sd)
sd_mpg_gear

```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!

cyli_gear_obser <- table(mtcars$cyl, mtcars$gear)
cyli_gear_obser 

print("The most common car type in this data set is car with __8__ cylinders and ___3_ gears. There are total of ___12__ cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
plot(weight ~ group, data = PlantGrowth, 
        main = "Plant Growth by Group", 
        xlab = "Group", 
        ylab = "Weight")
summary(PlantGrowth)

```

Result:

=> Report a paragraph to summarize your findings from the plot!

After creating a graph for PlantGrowth, I have found that ctrl had a median of 5.1. I had also found that they had a min of 3.5 and a max of 6.3. I had found that trt1 had a median of 4.5. I have also found that it had a min of 3.6 and a max of 5.8. Lastly I have found that trt2 had a mean of 5.5. I had also found that it had a min of 5.0 and a max of 6.4. Overall, I fit that ctr had a better mediam, min, and max. 

b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}

hist(mtcars$mpg, breaks = 10, 
     main = "Histogram of MPG", 
     xlab = "Miles per gallon", 
     ylab = "Num of Cars")

print("Most of the cars in this data set are in the class of ___15___ mile per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
pairs(USArrests, main = "Pairs Plot of USArrests Data")

plot(USArrests$Murder, USArrests$Assault, 
     main = "Scatter Plot of Murder and Assault", 
     xlab = "Murder", 
     ylab = "Assault")

```

Result:

=> Report a paragraph to summarize your findings from the plot!
From the pair, I have found that it gave a bunch of different plots of different of arrest. it was very interesting to see that all of the plots were very different from one another. 
As From the plot of arrest and murder, I have found that the plots are everywhere. However i do see a strong positive correlation between murder and arrest. I see that they both are both increasing together.

***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!

#descriptive statistics
summary(housingData)
summary(housingData$Year.Built)
summary(housingData$Market.Value.per.SqFt)

max(housingData$Market.Value.per.SqFt)
min(housingData$Market.Value.per.SqFt)

m <- mean(housingData$Market.Value.per.SqFt)
v <- var(housingData$Market.Value.per.SqFt)
s <- sd(housingData$Market.Value.per.SqFt)

print(paste("The average of Market Value Per SqFt from this data set is ", m , " with variance ", v , " and standard deviation", s , "."))

#aggregation tables 
avg_MVPSqFt_Year <- aggregate(Market.Value.per.SqFt ~ Year.Built, data = housingData, mean)
avg_MVPSqFt_Year

avg_MVPSqFt_Boro <- aggregate(Market.Value.per.SqFt ~ Boro, data = housingData, mean)
avg_MVPSqFt_Boro

avg_MVPSqFt_Fin <- aggregate(Market.Value.per.SqFt ~ Neighborhood, data = housingData, mean)
avg_MVPSqFt_Fin

sd__MVPSqFt_Year <- aggregate(Market.Value.per.SqFt ~ Year.Built, data = housingData, sd)
sd__MVPSqFt_Year

var__MVPSqFt_Year <- aggregate(Market.Value.per.SqFt ~ Year.Built, data = housingData, var)
var__MVPSqFt_Year


```


b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
hist(housingData$Market.Value.per.SqFt, breaks = 10,
     main = "Histogram of Market Value per SqFt", 
     xlab = "Market Value per SqFt", 
     ylab = "Frequency")

plot(avg_MVPSqFt_Year$Year.Built, avg_MVPSqFt_Year$Market.Value.per.SqFt,
     main = "Scatter Plot of Market Value per SqFt and the years", 
     xlab = "Year", 
     ylab = "Market Value per SqFt")

```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!
From this exercise, I have found that max of Market value per SqFt were 399.38 while the min is 10.66. This was shocking to see the range from the max to the min. I always thought the max would be bigger and the min would be smaller. I have found that the average of Market Value Per SqFt is  133.173098814229, while variance is 4744.94274286489  and standard deviation 68.8835447902102. The Summary of housingdata was interesting to see, the most shocking part was seeing that the min was 1825 and the max was 2010. Some of the aggregate I had created also showed interesting finds. One of aggregate i will be pointing was the most interesting one out of all of the aggregates. This aggregate showed what year the house was built and the average of the Market value per SqFt. It showed that in 1825 the Market value per SqFt was 76.36000	whereas in 2010 the Market value per SqFt was	90.36667. This helps shows that the Market value per SqFt has changed significantly throughout the years. Looking at the graph i had created. I found that both showed very interesting data however, I am going to talk about the most interesting graph which is the scatter plot of market value per SqFt and the years. I found it so interesting the the plots where kind of everywhere but at the same time it kind of stayed in the same area. The area the plot seemed to stay on was from 100 -200 market value per SqFt between the years of 19090 to 2000. There seemed to be no positive increase or negative decrease. It seem like the data stayed static. There is a very weak or almost no correlation. Overall, all of my findings were interesting and i enjoyed learning about this exercise.  

